IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-07-06 paper/retrofit.tex
First stab at swivel
Signed-off-by: Edward Z. Yang <ezyang@cs.stanford.edu>

@@

\subsection{Examples}

To get a feeling for how these operational semantics operate, we
describe a few sample programs, and their corresponding behavior.

\paragraph{Sending mutable references}  As each task maintains its own
global state for the target language, one might reasonably wonder what would
happen if you attempted to send a reference to another task.
\begin{align*}
    & |let i = TI (sandbox (brecvf x _ (IT (!(TI x)))))|\\
    & |in TI (send (IT i) getLabel (IT (ref true)))|
\end{align*}
It would be a violation of information flow if such an operation actually
resulted in references being shared between two tasks.  In fact, the address
will be received by the sandboxed task, but it will be a dangling pointer; the
task will get stuck (under the |roundrobinf| scheduler the \textsc{I-noStep} will in turn remove it from the task list).  Readers
who are curious whether or not a static discipline could be enforced to
prevent these dangling pointers should check Section~\ref{sec:extensions}.

\paragraph{Status of the heap after a sandbox}  Consider a program whose
semantics depends on the definition of |klone|:
\begin{align*}
    & |let x = ref true| \\
    & |in TI (sandbox (IT (!x; !x))); x := false|
\end{align*}
When |klone| is the identity, the sandboxed task successfully dereferences
the contents of the reference to be |true| twice (recall that references
are cloned into the new global state).  When |klone| is the constant function
to the empty store, |x| is a dangling pointer.

\paragraph{Exceptions}  One language feature that has caused difficulty for
many information flow control systems is handling exceptions.
%
The key difficulty appears when considering the propagation of exceptions
across stack boundaries where the current label is restored, e.g., from secret
to public, as done by LIO's |toLabeled| and Breeze's
brackets~\cite{Hritcu:2013:YIB:2497621.2498098, stefan:2012:arxiv-flexible}.
%
If the exception crosses this boundary, it can leak information about the
inner secret computation into a public computation; hence, both systems
explicitly ensure that |toLabeled| does not propagate exceptions.

Our general embedding approach provides these semantics by construction.
%
To show this, we first define an encoding of the sequential |toLabeled|
construct in our framework:
\begin{align*}
    & |toLabeled l te @= | \\
    & |tab let i = taskId in| \\
%%    & |tab let j = TI (sandbox (IT (TI (setLabel l); TI (send i l (IT te)))))| \\
    & |tab let j = TI (sandbox (IT (TI (send i l (IT te)))))| \\
    & |tab in (\ _ . TI (setLabel l); TI (recvf x _ x unit))|
\end{align*}
When using the sequential scheduler |seqf|, the sandboxed task runs to
completion before returning to the task that sandboxed it. 
(Here, a task list can be interpreted as a number of information flow aware
stack frames.)
To read the result of a |toLabeled| operation
(which is communicated via message passing), the parent task must raise
its label; otherwise the label will get dropped, as per the \textsc{I-recv}
rule.\footnote{A keen reader may note that with a single, per-task channel, we
might mix up our messages.  Indeed, a proper development requires the ability
to allocate fresh channels; for purposes of simplicity, this has been omitted.}

Now suppose that our |targetLangML| is extended with a single exception
expression |fail|, which reduces to |fail|, removing all containing |tE|
evaluation contexts, i.e., it propagates up to a |IT (-)| boundary.
%
In this setting, it is clear that exceptions cannot cross boundaries.
%
For instance, 
\begin{code}
TI(toLabeled l (IT (if te then fail else true))); true
\end{code}
always evaluates to |true|, regardless of the value of |te|.
%
Just like in LIO and Breeze, however, only when calling |recv| on such messages
will the exception be propagated.
%
This of course, is not a leak since the current label must be raised before
receiving sensitive message, regardless of their content.
\Red{TODO: note that if the target language allows inspecting if
  result is exception we can implement Breeze's NAV (vs. just
propagating)}

>>>>

Before delving into the formal operational semantics in earnest, we'd
like to first demonstrate some examples where a dynamic information flow
control system can either be thought of as the execution of a single
language or a collection of communicating, isolated tasks.

\Red{(need to describe the correspondence)}

\paragraph{Language restrictions versus isolated sandboxes}  If you want
to create a standard programming language with information-flow control,
the usual technique is to build in restrictions (dynamic or static)
which allow noninterference to be proved.  While this is fine for toy
languages, if the language in question is an existing language, these
modifications can often require extremely intrusive, both in terms of
semantic changes and implementation changes, and it is easy to get things
wrong and permit a subtle implicit flow between different security levels.

The alternative method is to start with complete isolation between
different security levels: literally seperate instances of the
``language runtime'' for each traditional IFC thread.  For languages
which already have isolated implementations, such as JavaScript and C
with access to hardware privilege separation, this is a far more
practical starting point.  From here, the line between the isolated
instances can be blurred, but the `completely isolated'' semantics
are a helpful guidepost for what can be done safely.

As an example, imagine a program which attempts to send a mutable
reference from a low security level to a high security level:
\begin{align*}
    & |let i = TI (sandbox (brecvf x _ (IT (!(TI x)))))|\\
    & |in TI (send (IT i) getLabel (IT (ref true)))|
\end{align*}

For now, ignore the |TI| and |IT| annotations: this code creates a new
sandboxed thread $i$ which waits for a message, and then sends the
thread a mutable cell containing the value true.  If this mutable cell
is actually shared between the two tasks, it would be a violation of
information flow if such an operation actually resulted in references
being shared between two tasks.  However, since these are isolated
tasks, the address received by the sandboxed task is actually a dangling
pointer.  We can use a static discipline to avoid this situation
entirely or dynamically copy the mutable cell into the destination
context; in any case, the isolation makes it clear what semantics can
be safely supported.

While this is a very simple example, the elaborations of this principle
can actually be quite non-obvious for more interesting language
features.

\paragraph{Stack frames versus scheduling}  A feature that is supported
by some information flow control systems is the ability to temporarily
raise one's label in order to do computation on data.  Once the
computation is finished, the label is lowered back to the original
label---a stack discipline.  Care must be taken to ensure that
information doesn't leak when the label is lowered (for example, the
parent is not allowed to inspect the result of the computation). In
fact, systems with this feature generally only achieve
termination-insensitive noninterference: if the high secrecy section
doesn't terminate, this is observable by the halted execution of the low
secrecy section.

In this paper, we will instead say that the allocation of a new
stack frame with a raised label corresponds to the creation of a
new isolated task; a sample encoding using our language is
shown in Figure~\ref{toLabeled}.

\begin{figure}
\begin{align*}
    & |toLabeled l te @= | \\
    & |tab let i = taskId in| \\
%%    & |tab let j = TI (sandbox (IT (TI (setLabel l); TI (send i l (IT te)))))| \\
    & |tab let j = TI (sandbox (IT (TI (send i l (IT te)))))| \\
    & |tab in (\ _ . TI (setLabel l); TI (recvf x _ x unit))|
\end{align*}
\caption{Encoding of toLabeled using isolated tasks.}\label{fig:toLabeled}
\end{figure}

The sequential processing of these
tasks corresponds to a specific \emph{scheduling policy} of these
threads.  This change of perspective to scheduling highlights the
essential differents between termination-insensitive and termination-sensitive
noninterference.

\paragraph{Exceptions versus message passing}
A language feature that has caused difficulty for
many information flow control systems is handling exceptions.
%
The key difficulty appears when considering the propagation of exceptions
across stack boundaries where the current label is restored, e.g., from secret
to public, as done by LIO's |toLabeled| and Breeze's
brackets~\cite{Hritcu:2013:YIB:2497621.2498098, stefan:2012:arxiv-flexible}.
%
If the exception crosses this boundary, it can leak information about the
inner secret computation into a public computation; hence, both systems
explicitly ensure that |toLabeled| does not propagate exceptions.

When stack frames are treated as isolated tasks, the correct semantics
are achieved automatically.
%
The only way for the stack frame to communicate information is by
initiating inter-task message passing; in its absence, the ``parent
frame'' will operate as if no exception had occurred.
%
For instance, 
\begin{code}
TI(toLabeled l (IT (if te then fail else true))); true
\end{code}
always evaluates to |true|, regardless of the value of |te|.
%
Of course, when communication is performed, there is no leak,
because all of the necessary information flow checks needed to
communicate between tasks are also performed here.
\Red{TODO: note that if the target language allows inspecting if
  result is exception we can implement Breeze's NAV (vs. just
propagating)}

\subsection{Preliminaries}

@@
